\subsubsection{External}
\label{External}
The \textbf{External} post-processor will execute an arbitrary python function
defined externally using the \textit{Functions} interface (see
Section~\ref{sec:functions} for more details).
%

\ppType{External}{External}

\begin{itemize}
  \item \xmlNode{method}, \xmlDesc{comma separated string, required field},
  lists the method names of an external Function that will be computed (each
  returning a post-processing value). \nb New variable names will be defined as:
  ``Function Name in this post-processor'' + ``\_`` + ``variable name in XML
  node \xmlNode{method}''. These new variables will be used to store the computed
  values from the list of methods, and can be accessed by the users through RAVEN
  entities \textbf{DataObjects} and \textbf{OutStreams}.
  \item \xmlNode{Function}, \xmlDesc{xml node, required string field}, specifies
  the name of a Function where the \textit{methods} listed above are defined.
  %
  \nb This name should match one of the Functions defined in the
  \xmlNode{Functions} block of the input file.
  %
  The objects must be listed with a rigorous syntax that, except for the XML
  node tag, is common among all the objects.
  %
  Each of these sub-nodes must contain 2 attributes that are used to map them
  within the simulation framework:

   \begin{itemize}
     \item \xmlAttr{class}, \xmlDesc{required string attribute}, is the main
     ``class'' the listed object is from, the only acceptable class for
     this post-processor is \xmlString{Functions};
     \item \xmlAttr{type}, \xmlDesc{required string attribute}, is the object
     identifier or sub-type, the only acceptable type for this post-processor is
     \xmlString{External}.
  \end{itemize}
\end{itemize}

  This Post-Processor accepts as Input/Output both \xmlString{PointSet} and \xmlString{HistorySet}:
   \begin{itemize}
    \item If a \xmlString{PointSet}  is used as Input, the parameters are passed in the external  \xmlString{Function}
  as numpy arrays. The methods' return type must be either a new array or a scalar. In the following it is reported an example
  with two methods, one that returns a scalar and the other one that returns an array:
      \begin{lstlisting}[language=python]
import numpy as np
def sum(self):
  return np.sum(self.aParameterInPointSet)

def sumTwoArraysAndReturnAnotherone(self):
  return self.aParamInPointSet1+self.aParamInPointSet2
      \end{lstlisting}
    \item If a \xmlString{HistorySet}  is used as Input, the parameters are passed in the external  \xmlString{Function}
     as a list of numpy arrays. The methods' return type must be either a new list of arrays (if the Output is another
     \xmlString{HistorySet}), a scalar or a single array (if the  Output is  \xmlString{PointSet} . In the following it
     is reported an example
     with two methods, one that returns a new list of arrays (Output = HistorySet) and the other one that returns an array (Output =
     PointSet):
      \begin{lstlisting}[language=python]
import numpy as np
def newHistorySetParameter(self):
  x = []*len(self.time)
  for history in range(len(self.time)):
    for ts in range(len(self.time[history])):
      if self.time[history][ts] >= 0.001: break
    x[history] = self.x[history][ts:]
  return x

def aNewPointSetParameter(self):
  x = []*len(self.time)
  for history in range(len(self.time)):
    x[history] = self.x[history][-1]
  return x
      \end{lstlisting}
   \end{itemize}

\textbf{Example:}
\begin{lstlisting}[style=XML,morekeywords={subType,debug,name,class,type}]
<Simulation>
  ...
  <Models>
    ...
    <PostProcessor name="externalPP" subType='External' verbosity='debug'>
      <method>Delta,Sum</method>
      <Function class='Functions' type='External'>operators</Function>
        <!-- Here, you can add a Function defined in the
             Functions block. This should be present or
             else RAVEN will not know where to find the
             defined methods. -->
    </PostProcessor>
    ...
  </Models>
  ...
</Simulation>
\end{lstlisting}

\nb The calculation results from this post-processor are stored in the internal variables. These variables
are accessible by the users through RAVEN entities \textbf{DataObjects} and \textbf{OutStreams}. The names
of these variables are defined as: ``Function Name in this post-processor'' + ``\_`` + ``variable name in XML
node \xmlNode{method}''. For example, in previous case, variables ``operators\_Delta'' and ``operators\_Sum''
are defined by RAVEN to store the outputs of this post-processor.
