%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% SERPENT  INTERFACE  %%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{SERPENT Interface}
\label{subsec:serpentInterface}
The Serpent interface is meant to run multiple user-defined SERPENT simulations
while storing the values contained in the following output file types: 
\begin{itemize}
\item Results' file (``\_res.m'')
\item Detectors' files  (``\_det[bu].m'')
\item Depletion file (``\_dep.m'')
\item Burn-up matrices' files (``\_depmtx\_[mat]\_[bu]\_[ss].m'')
\end{itemize}

where:
\begin{itemize}
\item  \texttt{bu} is the burn up step index
\item  \texttt{mat} is the material name
\item  \texttt{ss} is the sub step id (if depletion with sub-steps is activated).
\end{itemize}

The interface allows users to run SERPENT by varying values from a 
template input file using the wildcard approach available through the
generic code interface. In section \ref{subsec:genericInterface}  information
on how to use the wildcard approach and all the available options of the interface
can be found.

\subsubsection{Models}
\label{subsubsec:serpentInterfaceModels}
In order to run the code, make sure you have a valid SERPENT input file.

In the RAVEN input file, your \xmlNode{Models} node should look like:
\begin{lstlisting}[style=XML]
<Models>
    <Code name="SERPENT" subType="Serpent">
      <!-- path to your serpent executable -->
      <executable>/my/path/to/serpent/sss2</executable>
      <clargs arg="" extension=".serpent" type="input"/>
      <clargs arg="--noplot" type="postpend"/>
      <additionalFileTypes>
        DepletionReader, DetectorReader, DepmtxReader
      <additionalFileTypes/>
      <!-- EOL parameters -->
      <EOL target="absKeff">1.0</EOL>
      <EOL target="impKeff">1.0</EOL>
    </Code>
</Models>
\end{lstlisting}
where the \xmlNode{executable} and \xmlNode{clargs} should me modified
to create the appropriate run command for SERPENT. In this example, the command
to run raven is:
\begin{lstlisting}[language=bash]
/my/path/to/serpent/sss2 [inputFile] --noplot
\end{lstlisting}
where the \texttt{inputFile} is defined in the \xmlNode{Files} node as:
\begin{lstlisting}[style=XML]
<Files>
    <Input name="originalInput" type="">serpent_input.serpent</Input>
</Files>
\end{lstlisting}

Additional XML sub-nodes can be optionally added in the \xmlNode{Code} XML node is:
\begin{itemize}
   \item \xmlNode{additionalFileTypes}, \xmlDesc{comma separated list, optional parameter},  additional output file types 
                                                              that the interface needs to process.  
                                                              Available options are:  
                                                              \begin{itemize}
                                                                \item  \textit{ResultsReader} (``*\_res.m''), Results’ file.
                                                                \item  \textit{DepletionReader} (``*\_dep.m''), Depletion file.
                                                                 \item \textit{DetectorReader}  (``*\_det[bu].m''), Detectors’ files.
                                                                \item  \textit{DepmtxReader} (``*\_depmtx\_[mat]\_[bu]\_[ss].m''), Burn-up matrices’ files.
                                                              \end{itemize}
                                                              \nb By default  the ``Results'' (``\_res.m'') file (using the ``ResultReader'') is the 
                                                               only output file processed (``*\_res.m'') by the interface. 
                                                               For variable processing conventions, see section  
                                                               \ref{subsubsec:serpentInterfaceOutputConversion}.
   \item \xmlNode{EOL}, \xmlDesc{float, optional parameter},  value of the ``target'' at which the time (in burnup calculations) of end of life should be recorded.
                                                             The  \xmlNode{EOL} node  must contain the attribute \xmlAttr{target}, which indicates the FOM in the Results file that should be
                                                             monitored (e.g. ``absKeff''). The user can specify as many  \xmlNode{EOL} as needed.
                                                              
                                                              \nb  The output parser will generate an output variable called ``EOL\_\textit{target}'' that represents the time at which the 
                                                              parameter ``target'' takes the value indicated in this node. For example,  if ``target'' == absKeff and the value of the XML node is 1.0, 
                                                              the variable ``EOL\_absKeff'' will contain the time (burnDays) at which the ``absKeff'' == 1.0.
                                                              
\end{itemize}

\subsubsection{Files}
The only input file needed is a complete SERPENT input file,
which means that it should either be self sufficient, or includes
the necessary files (e.g. geometry, material definition files).
The main \textit{SERPENT} input file must be tagged with the
type ``serpent'' (as shown below). The auxiliary files, included in the
main input file through the SERPENT include nomenclature (i.e., `` include "myAuxFile" ''),
do not need to be tagged with any specific type:

\begin{lstlisting}[style=XML]
  <Files>
    <Input name="sMainInput" type="serpent">
      mySerpentInputFile.serpent
     </Input>
    <Input name="detectors"  type="">
     detectors
     </Input>
    <Input name="geometry"  type="">
      geometry
    </Input>
  </Files>
\end{lstlisting}

\nb Since the Generic Code interface (see section  \ref{subsec:genericInterface})
      is used to perturb the SERPENT input file, the wild-carded variables to perturb
      can be present in any of the files needed by SERPENT (main input file and its ``include'' files).

\subsubsection{Samplers / Optimizers}
In the \xmlNode{Samplers} block the user can define the variables
to be sampled.
\textbf{Example:} If the user wants to vary depletion time from
10 to 100 days, the SERPENT input file \texttt{dep} definition is as such:
\begin{lstlisting}
dep daystep 
$RAVEN-deptime$ 
\end{lstlisting}

Then in the RAVEN input file, the \xmlNode{Samplers} block would be:
\begin{lstlisting}[style=XML]
  <Samplers>
    <Grid name="myGrid">
      <variable name="deptime">
        <distribution>timedist</distribution>
        <!-- equally spaced steps with lower and upper bound -->
        <grid construction="equal" steps="100" type="CDF">0.1 1</grid>
      </variable>
    </Grid>
  </Samplers>
\end{lstlisting}
where the distribution \texttt{timedist} is defined in the \xmlNode{Distributions}:
\begin{lstlisting}[style=XML]
  <Distributions>
    <Uniform name="timedist">
      <lowerBound>0.0</lowerBound>
      <upperBound>100</upperBound>
    </Uniform>
</Distributions>
\end{lstlisting}

Then this run would be defined in the \xmlNode{Steps} block as \xmlNode{MultiRun}:
\begin{lstlisting}[style=XML]
<MultiRun name="runGrid">
      <!-- runGrid runs serpent by 
           the number of steps with
           sampled variable 
       -->
      <Input class="Files" type="serpent">
        sMainInput
      </Input>
      <Input  class="Files" type="">
        detectors
      </Input>
      <Input class="Files" type="">
        geometry
      </Input>
      <Model class="Models" type="Code">
        SERPENT
      </Model>
      <Sampler class="Samplers" type="Grid" >
        myGrid
      </Sampler>
      <Output class="DataObjects" type="PointSet">
        outPointSet
      </Output>
</MultiRun>
\end{lstlisting}

\subsubsection{Output Files Conversion}
\label{subsubsec:serpentInterfaceOutputConversion}
SERPENT can generate several output files. The description of 
such output files can be found in SERPENT documentation page  
(available at \url{https://serpent.vtt.fi/mediawiki/index.php/Main_Page}).

As mentioned in previous sections, the interface processes, by default, the
main Results file \texttt{[input]\_res.m} (both in time-independent  and burnup calculations).
Additional output files can be requested, as described in Section 
\ref{subsubsec:serpentInterfaceModels}. Such output types are:
\begin{itemize}
   \item  \textit{ResultsReader} (``*\_res.m''), Results’ file.
   \item  \textit{DepletionReader} (``*\_dep.m''), Depletion file.
   \item \textit{DetectorReader}  (``*\_det[bu].m''), Detectors’ files.
   \item  \textit{DepmtxReader} (``*\_depmtx\_[mat]\_[bu]\_[ss].m''), Burn-up matrices’ files.
\end{itemize}

In this section, the naming conventions for the variables in each of the available output types 
are reported.\\

\textit{\textbf{\underline{ResultsReader}}}
\\The \textit{ResultsReader} is the default output parsing option. 
The reader is responsible to read the result files (``*\_res.m'').
The parser reads all the results (time-independent  or burnup dependent) that can
be found in the file.
The variable names are converted from using underscores (``\_'') to \textit{\textbf{camelBack}}. For example,
the variables \textit{\textbf{ANA\_KEFF}} and  \textit{\textbf{IMP\_KEFF}}  will be renamed
 as \textit{\textbf{anaKeff}} and \textit{\textbf{impKeff}}, respectively. 
 \\In addition, the variables characterized by multiple elements will be expanded in unique variables, 
 appending to the variable name the element index (e.g. 0, 1, 2, etc.). 
 \begin{lstlisting}[language={matlab}]
% _res.m file
ANA_KEFF  (idx, [1:   6]) = [1.0E+00 0.001 9.9E-01 0.001 1.0E-02 0.001];
IMP_KEFF  (idx, [1:   2]) = [1.0E+00 0.001];
\end{lstlisting}
For example, for the above snippet of the result file, the variables generated would be:

\begin{table}[ht]
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\textit{anaKeff\_0} & \textit{anaKeff\_1} & \textit{anaKeff\_2} & \textit{anaKeff\_3} & \textit{anaKeff\_4} & \textit{anaKeff\_5} & \textit{impKeff\_0} & \textit{impKeff\_1} \\ \hline
1.E+00              & 0.001               & 9.9E-01             & 0.001               & 1.E-02              & 0.001               & 1.E+00              & 0.001               \\ \hline
\end{tabular}
\end{table}

For burnup calculations the same syntax is used with the different that multiple rows are generated each of them indexed by the burnup step.

\textit{\textbf{\underline{DepletionReader}}}
\\The \textit{DepletionReader} is aimed to parse and process depletion files.
The reader is responsible to read the depletion file (``*\_dep.m'').
The parser reads all the depletion results that can
be found in the file.
The variable names are converted from using underscores (``\_'') to \textit{\textbf{camelBack}}. For example,
the variable \textit{\textbf{ING\_TOX}} will be renamed
 as \textit{\textbf{ingTox}}.
 The parser parses the materials, isotopes and quantities for creating unique variables, based on the following nomenclature:
  \begin{lstlisting}[language={matlab}]
   [materialName]_[isotope]_[quantity]
\end{lstlisting}
For example, The variable referring to the  \textit{activity}  of the  \textit{U235}  isotope in material  \textit{UO2} will be called:
\textit{UO2\_U235\_activity}.
For global variables (not linked to specific isotopes) the same nomenclature reported above is used removing the ``isotope'' section.
For example, the burnup of the material \textit{UO2} will be named \textit{UO2\_burnup}.

\textit{\textbf{\underline{DetectorReader}}}
\\The \textit{DetectorReader} is aimed to parse and process the detector output files.
The reader is responsible to read the detector files (``*\_det[bu].m'') for both time-independent and burnup calculations.
The parser reads all the detector results that can
be found in the files.
For detector returning scalar quantities (E.g. total flux on a surface), the variable nomenclature is as follows:
  \begin{lstlisting}[language={matlab}]
   for the actual scalar result:
   [dName]
   for the uncertainty/error associated with the detector score:
   [dName]_err
\end{lstlisting}
For detector returning a field quantity (e.g. energy and spatial dependent flux), the variable nomenclature is as follows:
  \begin{lstlisting}[language={matlab}]
   for the actual detector score entry
   [dName]_[gridIndexName1]_[gridIndexValue1] (...)_[gridIndexNameN]_[gridIndexValueN]  ]
   for the uncertainty/error associated with the detector score entry:
   [dName]_[gridIndexName1]_[gridIndexValue1] (...)_[gridIndexNameN]_[gridIndexValueN]_err
\end{lstlisting}


\textit{\textbf{\underline{DepmtxReader}}}
\\The \textit{DepmtxReader} is aimed to parse and process Burn-up matrices’ files.
Since the matrices can be quite large, the parser reads the matrices and store them in a serialized file.
The filename is returned to RAVEN following the nomenclature:
  \begin{lstlisting}[language={matlab}]
    filename_depmtx_zai_[materialName]_[bu]_[ss]
\end{lstlisting}
 In addition, the 1 group flux used for the depletion is returned as follows:
  \begin{lstlisting}[language={matlab}]
     flx_[materialName]_[bu]_[ss]
\end{lstlisting} 

\subsubsection{SERPENT/RAVEN variable generation through script}
Since the number of output variables generated by SERPENT and processable by SERPENT/RAVEN interface is quite large and,
consequentially, manually listing them would be cumbersome, 
an utility script has been provided to create  \xmlNode{VariableGroups} (\xmlNode{Group}s XML nodes) containing all the
variables (with the correct nomenclature as seen in section \ref{subsubsec:serpentInterfaceOutputConversion} extractable by
specific output files.
\\In order to link the variable generation to the specific SERPENT case of interest, the script inspects pre-generated SERPENT output files
provided by the user.
\\The script is located at \path{RAVEN_HOME/ravenframework/CodeInterfaceClasses/SERPENT/utilityForCreatingVariables/generateRavenVariables.py} and
can be invoked like indicated (for example) below:
\begin{lstlisting}[language=bash]
python generateRavenVariables.py -fileTypes ResultsReader,DepletionReader,DetectorReader,DepmtxReader   -fileRoot mySerpentInputFile -o myOptionalScriptOutputFilename
\end{lstlisting}
where:
\begin{itemize}
\item  \texttt{-fileTypes}, Comma-separated list of output file types from which variable names need to be extracted. 
                                     Currently, the available types  are the ones indicated in \ref{subsubsec:serpentInterfaceOutputConversion} 
                                     (e.g. ResultsReader,DepletionReader,DetectorReader,DepmtxReader)
\item  \texttt{-fileRoot}, File name root from which all the output files are going to be inferred. 
                                    For example, if the ``fileRoot'' is ``myInputFile'', the ``ResultsReader'' file that the script will parse
                                    must be named ``myInputFile\_res.m''.
\item  \texttt{-o}, Optional output file name. If not provided, the script will store the  \xmlNode{VariableGroups} in file named \path{ravenOutputVariables.xml}.
\end{itemize}


