\section{Introduction}

% High-level RAVEN description
RAVEN is a generic software framework to perform parametric and probabilistic
analysis based on the response of complex system codes.
%
The initial development was aimed at providing dynamic risk analysis
capabilities to the thermohydraulic code RELAP-7, currently under development
at Idaho National Laboratory (INL).
%
Although the initial goal has been fully accomplished, RAVEN is now a
multi-purpose probabilistic and uncertainty quantification platform, capable of
agnostically communicating with any system code.

% RAVEN's flexibility
The agnosticism of RAVEN is inherent in its provision of Application Programming
Interfaces (APIs).
%
These APIs allow RAVEN to interact with any code as long as all the parameters
that need to be perturbed are accessible by input files or via python
interfaces.
%
%RAVEN is capable of investigating the system response, and investigating the
%input space using Monte Carlo, Grid, or Latin Hyper Cube sampling schemes, but
%its strength is focused towards system feature discovery, such as limit
%surfaces, separating regions of the input space leading to system failure, and
%using dynamic supervised learning techniques.
RAVEN is capable of investigating system response and input space using various
sampling schemes such as Monte Carlo, grid, or Latin hypercube.
%
However, its strength lies in system feature discovery such as: constructing
limit surfaces, separating regions of the input space leading to system failure,
and using dynamic supervised learning techniques.

% Historical Context
The development of RAVEN started in 2012, when, within the Nuclear Energy
Advanced Modeling and Simulation (NEAMS) program, the need to provide a modern
risk evaluation framework arose.
%
RAVEN's principal assignment is to provide the necessary software and algorithms
in order to employ the concepts developed by the Risk Informed Safety Margin
Characterization (RISMC) program.
%
RISMC is one of the pathways defined within the Light Water Reactor
Sustainability (LWRS) program.

% RISMC
In the RISMC approach, the goal is not just to identify the frequency of an
event potentially leading to a system failure, but the proximity (or lack
thereof) to key safety-related events.
%
Hence, the approach is interested in identifying and increasing the safety
margins related to those events.
%
A safety margin is a numerical value quantifying the probability that a safety
metric (e.g. peak pressure in a pipe) is exceeded under certain conditions.

% Conclusion
The initial development of RAVEN has been focused on providing dynamic risk
assessment capability to RELAP-7, currently under development at INL.
%
Most of the capabilities, implemented having RELAP-7 as a principal focus, are
easily deployable to other system codes.
%
For this reason, several side activates have been employed (e.g.  RELAP5-3D, any MOOSE-based App, etc.)
or are currently ongoing for coupling RAVEN with several different software.
%
The aim of this document is to detail the input requirements for RAVEN focusing
on the input structure.
