 \section{RAVEN:SYSTEM REQUIREMENTS} 
 \subsection{Requirements Traceability Matrix} 
 This section contains all of the requirements, requirements' description, and 
 requirement test cases. The requirement tests are automatically tested for each 
 CR (Change Request) by the CIS (Continuous Integration System). 
 \newcolumntype{b}{X} 
 \newcolumntype{s}{>{\hsize=.5\hsize}X} 
 \subsubsection{Minimum Requirements} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-M-1 & \hspace{0pt}Computer: Any POSIX (and POSIX-like) system & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
 \hspace{0pt}R-M-2 & \hspace{0pt}RAM: 2 GB per core execution (depending on the type of analysis and data genarated) & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
 \hspace{0pt}R-M-3 & \hspace{0pt}Disk: 10 GB (size depending on the type of analysis and data generated) & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
 \hspace{0pt}R-M-4 & \hspace{0pt}Compilers: GCC, Clang, or Intel & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
 \hspace{0pt}R-M-5 & \hspace{0pt}Language: Python 2.7 & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
 \hspace{0pt}R-M-6 & \hspace{0pt}Version Control: Git & \hspace{0pt}1)``RAVEN User Manual'', INL/EXT-15-34123 2)Continous Integration System \\ \hline 
\hline 
\caption*{Minimum Requirements}
\end{tabularx} 
 \subsubsection{Functional Requirements} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-F-1 & \hspace{0pt}RAVEN shall allow support for user-defined instructions for controlling the execution stages of a simulation. & \hspace{0pt}1)/raven/tests/framework/test\_rom\_trainer.xml 2)/raven/tests/framework/test\_random.xml \\ \hline 
\hline 
 \hspace{0pt}R-F-2 & \hspace{0pt}RAVEN shall allow for user-defined resource allocation for driving external applications. & \hspace{0pt}1)/raven/tests/framework/CodeInterfaceTests/generic\_parallel.xml \\ \hline 
\hline 
 \hspace{0pt}R-F-3 & \hspace{0pt}RAVEN shall support a programmatic method for building up and/or downloading the necessary compiled objects/dependencies necessary for a simulation. & \hspace{0pt}1)RAVEN User Manual, INL/EXT-15-34123 \\ \hline 
\hline 
 \hspace{0pt}R-F-4 & \hspace{0pt}RAVEN shall provide the ability to resume a previous simulation using data generated and exported by RAVEN itself. & \hspace{0pt}1)/raven/tests/framework/Samplers/Restart/test\_restart\_MC.xml 2)/raven/tests/framework/Samplers/Restart/test\_restart\_csv.xml 3)/raven/tests/framework/Samplers/Restart/test\_restart\_constant.xml \\ \hline 
\hline 
 \hspace{0pt}R-F-5 & \hspace{0pt}RAVEN shall allow for user-defined output types for simulation data. & \hspace{0pt}1)/raven/tests/framework/test\_output.xml 2)/raven/tests/framework/ROM/TimeSeries/DMD/test\_traditional\_dmd.xml \\ \hline 
\hline 
 \hspace{0pt}R-F-6 & \hspace{0pt}RAVEN shall allow for a standardized method for importing simulation data not previosly generated by the system itself. & \hspace{0pt}1)/raven/tests/framework/test\_output.xml 2)/raven/tests/framework/test\_iostep\_load.xml 3)/raven/tests/framework/Databases/test\_load\_and\_push\_reusing\_same\_hdf5.xml \\ \hline 
\hline 
\caption*{Framework, I/O, Execution Control}
\end{tabularx} 
 \subsubsection{Usability Requirements} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-RE-1 & \hspace{0pt}RAVEN shall support 1-Dimensional probability distributions including generating random numbers from them. & \hspace{0pt}1)/raven/tests/framework/unit\_tests/Distributions/TestDistributions.py \\ \hline 
\hline 
 \hspace{0pt}R-RE-2 & \hspace{0pt}RAVEN shall support N-Dimensional probability distributions. It shall support multivariate normal distributions and distributions defined by tabular data. & \hspace{0pt}1)/raven/tests/framework/test\_simple\_ND\_external\_MC.xml \\ \hline 
\hline 
 \hspace{0pt}R-RE-3 & \hspace{0pt}RAVEN shall support a variety of samplers that use probability distributions to sample the input space. & \hspace{0pt}1)/raven/tests/framework/test\_Grid\_Sampler.xml \\ \hline 
\hline 
\caption*{Risk Evaluation}
\end{tabularx} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-RA-1 & \hspace{0pt}RAVEN shall support adaptive sampling that use already gathered samples to determine where to locate new samples. & \hspace{0pt}1)/raven/tests/framework/PostProcessors/LimitSurface/test\_LimitSurface.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-2 & \hspace{0pt}RAVEN shall support importing/exporting data in CSV format. & \hspace{0pt}1)/raven/tests/framework/test\_iostep\_load.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-3 & \hspace{0pt}RAVEN shall support generating plots from the data it generates. & \hspace{0pt}1)/raven/tests/framework/test\_output.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-4 & \hspace{0pt}RAVEN shall be able to generate Reduced Order Models from its data and use them to predict responses from a system. & \hspace{0pt}1)/raven/tests/framework/test\_rom\_trainer.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-5 & \hspace{0pt}RAVEN shall be able to perform basic statistical analysis of generated data. & \hspace{0pt}1)/raven/tests/framework/PostProcessors/BasicStatistics/test\_BasicStatistics.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-6 & \hspace{0pt}RAVEN shall be able to perform advanced post processing of generated data, using data mining methodologies. & \hspace{0pt}1)/raven/tests/framework/PostProcessors/DataMiningPostProcessor/DimensionalityReduction/test\_dataMiningExactPCA.xml \\ \hline 
\hline 
 \hspace{0pt}R-RA-7 & \hspace{0pt}RAVEN shall be able to compute probability of failure based on generated data and goal functions & \hspace{0pt}1)/raven/tests/framework/PostProcessors/LimitSurface/test\_LimitSurface.xml 2)/raven/tests/framework/PostProcessors/LimitSurface/test\_LimitSurface\_and\_integral.xml \\ \hline 
\hline 
\caption*{Risk Analysis}
\end{tabularx} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-RM-1 & \hspace{0pt}RAVEN shall be able to choose the values of a set of input parameters that minimize/maximize a goal function that depends on system output figure of merits and input parameters. & \hspace{0pt}1)/raven/tests/framework/Optimizers/beale.xml \\ \hline 
\hline 
\caption*{Risk Mitigation}
\end{tabularx} 
 \subsubsection{Performance Requirements} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-IS-1 & \hspace{0pt}RAVEN shall be able to parallelize running external codes. & \hspace{0pt}1)/raven/tests/framework/CodeInterfaceTests/test\_LHS\_Sampler\_Bison\_parallel.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-2 & \hspace{0pt}RAVEN shall be able to run external codes by supplying them with the needed input files and collecting the output data. & \hspace{0pt}1)/raven/tests/framework/test\_simple.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-3 & \hspace{0pt}RAVEN shall support storing and retrieving data in a HDF5 database. & \hspace{0pt}1)/raven/tests/framework/Databases/test\_2steps\_same\_db.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-4 & \hspace{0pt}RAVEN shall be able to provide data to a user provided python function, and retrieve the data from that. & \hspace{0pt}1)/raven/tests/framework/test\_Lorentz.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-5 & \hspace{0pt}RAVEN shall be able to perform various calculation tasks (simulation and post processing), and transfer data to the next task. & \hspace{0pt}1)/raven/tests/framework/test\_calc\_and\_transfer.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-6 & \hspace{0pt}RAVEN shall be able to run external codes in parallel on shared memory machines. & \hspace{0pt}1)/raven/tests/framework/test\_bison\_mc\_simple.xml 2)/raven/tests/framework/CodeInterfaceTests/test\_generic\_interface.xml 3)/raven/tests/framework/CodeInterfaceTests/test\_generic\_interface\_custom\_out\_file.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-7 & \hspace{0pt}RAVEN shall be able to run external codes in parallel on distributed memory machines. & \hspace{0pt}1)raven/cluster\_tests/test\_mpi.xml 2)raven/cluster\_tests/test\_mpiqsub\_local.xml 3)raven/cluster\_tests/test\_pbs.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-8 & \hspace{0pt}RAVEN shall be able to run internal models in parallel on shared memory machines. & \hspace{0pt}1)/raven/tests/framework/InternalParallelTests/test\_internal\_parallel\_ROM\_scikit.xml 2)/raven/tests/framework/InternalParallelTests/test\_internal\_parallel\_extModel.xml 3)/raven/tests/framework/InternalParallelTests/test\_internal\_parallel\_PP\_LS.xml \\ \hline 
\hline 
 \hspace{0pt}R-IS-9 & \hspace{0pt}RAVEN shall be able to run internal models in parallel on distributed memory machines. & \hspace{0pt}1)raven/cluster\_tests/InternalParallel/test\_internal\_parallel\_extModel.xml 2)raven/cluster\_tests/InternalParallel/test\_internal\_parallel\_PP\_LS.xml 3)raven/cluster\_tests/InternalParallel/test\_internal\_parallel\_ROM\_scikit.xml \\ \hline 
\hline 
\caption*{Infrastructure Support}
\end{tabularx} 
 \subsubsection{System Interfaces} 
\begin{tabularx}{\textwidth}{|s|s|b|} 
\hline 
\textbf{Requirment ID} & \textbf{Requirment Description} & \textbf{Test(s)}  \\ \hline 
\hline 
 \hspace{0pt}R-SI-1 & \hspace{0pt}RAVEN shall be able to be coupled with external applications via input files. & \hspace{0pt}1)/raven/tests/framework/CodeInterfaceTests/RAVEN/rom.xml 2)/raven/tests/framework/CodeInterfaceTests/RELAP5/test\_relap5\_code\_interface.xml \\ \hline 
\hline 
 \hspace{0pt}R-SI-2 & \hspace{0pt}RAVEN shall be able to be coupled with external applications via Python API. & \hspace{0pt}1)/raven/plugins/ExamplePlugin/tests/test\_example\_plugin.xml 2)/raven/plugins/ExamplePlugin/tests/test\_raven\_running\_raven\_plugin.xml 3)/raven/tests/plugins/ExamplePlugin/test\_example\_plugin.xml 4)/raven/tests/plugins/ExamplePlugin/test\_raven\_running\_raven\_plugin.xml \\ \hline 
\hline 
\caption*{Interface with external applications}
\end{tabularx} 
\end{document}