\label{sec:introduction}
RAVEN (\textbf{R}eactor \textbf{A}nalysis and \textbf{V}irtual control \textbf{EN}viroment)~\cite{ravenFY12,mandelliANS2012} is a software tool that acts as the control logic driver for the newly developed Thermal-Hydraulic code RELAP-7  (\textbf{R}eactor \textbf{E}xcursion and \textbf{L}eak \textbf{A}nalysis \textbf{P}rogram). The goal of this report is to highlight the newly developed  Dynamic Event Tree module embedded in the code and its utilization in conjunction with RELAP-7. 

As for all PRA software the capability to control the plant evolution during the simulation itself is a plus as it is for the analysis of the propagation of the uncertainty. For these reasons, a strict interaction between RELAP-7 and RAVEN is a key of the long-term success of the overall project. In system safety analysis code a similar need is expressed by the implementation of the control logic of the plant. As a consequence the optimization of resources imposes the integration of this task under a common project that is naturally RAVEN. For this reason the control logic of the plant is simulated under RAVEN. This also offers the flexibility to easily implement proprietary control logic  without any change in the RELAP-7 code that is also regarded as a facilitating factor for the quick deployment of RELAP-7.

In order to summarize what RAVEN is, it can be said that it is a multi-purpose \textbf{P}robabilistic \textbf{R}isk \textbf{A}ssessment (PRA) software framework that allows dispatching different functionalities. 
It is designed to derive and actuate the control logic required to simulate the plant control system and operator actions (guided procedures) and to perform both Monte-Carlo sampling of random distributed events and Event Tree based analysis. 
In order to facilitate the input/output handling, a \textbf{G}raphical \textbf{U}ser \textbf{I}nterface (GUI) and a post-processing data mining module (under development) are available.

This report provides an overview of the DET structure, highlighting the mathematical framework from which its structure is derived and its software layout. In addition a \textbf{S}tation \textbf{B}lack \textbf{O}ut (SBO) DET based analysis of a simplified \textbf{P}ressurized \textbf{W}ater \textbf{R}eactor (PWR) model will be shown.
\vspace{-5mm}

%\subsection{subsection}
%text

%figure template

%\subsubsection{subsubsection}
%more text
%\paragraph{paragraph} 
%lot of text
%\subparagraph{subparagraph} 
%if you arrive at this point you have issues
