\subsection{System Operations}
\subsubsection{Human System Integration Requirements}
The command line interface shall support the ability to toggle any supported coloring schemes on or off pursuant to section 
508 of the Rehabilitation Act of 1973.
\subsubsection{Maintainability}
\begin{itemize}
  \item The latest working version (defined as the version that passes all tests in the current regression test suite) shall be 
           publicly available at all times through the repository host provider.
  \item  Flaws identified in the system shall be reported and tracked in a ticket or issue based system. The technical lead or 
            any COB member will 
            determine the severity and priority of all reported issues. The technical lead will assign resources at his or her 
            discretion to resolve identified issues.
  \item  The software maintainers will entertain all proposed changes to the system in a timely manner 
           (within two business days).        
  \item  The RAVEN software in its entirety will be made publicly available under the Apache version 2.0 license.     
\end{itemize}
\subsubsection{Human System Integration Requirements}
The regression test suite will cover at least 80\% of all lines of code at all times. 
The results of the regression tests will be stored in the Continuous Integration System.

\subsection{Information Management}
The RAVEN software in its entirety will be made publicly available on an appropriate repository hosting site (e.g. GitHub).
Backups and security services will be provided by the hosting service.

\section{Verification}
The regression test suite shall employ several verification tests using comparison against analytic 
solutions (when possible) and convergence rate analysis. 